---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Data is gathered from activity.csv and dplyr library is also imported
```{r}
library(dplyr)
data <-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
Total number of steps per date is calculated
```{r}
steps_byDate<-summarise(group_by(data,date),steps=sum(steps))
```
Histogram is created from the total number of steps taken per day
```{r}
hist(steps_byDate$steps,xlab="steps",main="histogram of total number of steps per day")
```
```{r , echo=FALSE}
mean<- mean(steps_byDate$steps,na.rm=TRUE)
median<- median(steps_byDate$steps,na.rm=TRUE)
```
Mean is `r mean`  
Median is `r median`

## What is the average daily activity pattern?
Total number of steps per interval is calculated
```{r}
steps_byInterval<-summarise(group_by(data,interval),steps=mean(steps,na.rm=TRUE))
```
Plot is created from the total number of steps taken per interval
```{r}
with(steps_byInterval,plot(x=interval,y=steps,type="l"))
```
```{r , echo=FALSE}
max<-subset(steps_byInterval,steps==max(steps),na.rm=TRUE)[,"interval"]
```
Max is `r max`  



## Imputing missing values
```{r, echo=FALSE}
missing_values<-sum(is.na(data))
```
There are `r missing_values` missing values  
Missing values are replaced with the mean
```{r}
data[is.na(data)]<-mean
steps_byDate1<-summarise(group_by(data,date),steps=sum(steps))
head(steps_byDate1)
```
Histogram is created with the new data set with replaced missing values
```{r }
hist(steps_byDate1$steps,xlab="steps",main="histogram of total number of steps per day")
```
```{r , echo=FALSE}
mean<- mean(steps_byDate1$steps)
median<- median(steps_byDate1$steps)
```
Mean is `r mean`  
Median is `r median` 

## Are there differences in activity patterns between weekdays and weekends?
New column days is created that is of type factor and shows whether a date is a weekend or a weekday
```{r }
weekend<- c("Saturday","Sunday")
data$days<-factor(weekdays(as.Date(data$date)) %in% weekend,levels=c(TRUE,FALSE),labels=c('weekend','weekday'))
head(data)
```
Number of steps per interval for weekends and weekdays are plotted
```{r }
steps_byDay<-summarise(group_by(data,interval,days),steps=mean(steps))
par(mfrow=c(1,2))
with(steps_byDay[steps_byDay$days=="weekend",],plot(x=interval,y=steps,type="l",col="red",main="weekend"))
with(steps_byDay[steps_byDay$days=="weekday",],plot(x=interval,y=steps,type="l",col="blue",main="weekday"))
```
